Static Race Detection in OS Kernels by Mining Locking Rules

### 2023-07-25 3.technique.tex
update technique

Given an alias graph, whether a variable and a lock exist in the same data 
structure can be determined by finding a common ancestor. If they are in the 
same data structure, the lock is likely to protect the variable when it is 
accessed. Figure~\ref{fig_pseudocode_get_access} shows the pseudocode to 
get the field of the accessed variable and the field of the protecting lock in 
the form of access path, if they exist in the same data structure. Given a node 
of an accessed variable and a node of a lock variable, the analysis first gets 
the common ancestor of the two nodes (Line 1). And then, if the common ancestor 
does not exist, the analysis returns NULL (Lines 2-3). Otherwise, the analysis 
gets the access path for the node of the accessed variable and the node of the 
lock variable from the common ancestor (Lines 5-7).

%\begin{table}[htbp]
%	\tablecaption{Notation table of pseudocodes.}
%	\label{tbl_notation}
%	\centering
%	\scriptsize
%	\begin{spacing}{0.98}
%		\begin{tabular}{@{ }p{2.7cm}p{5.05cm}@{ }}
%		\toprule[1pt]
%		GetAliasNode($\mathit{var}$, $\mathit{alias\_graph}$) & Get the node 
%		the variable $\mathit{var}$ exists in from the alias graph 
%		$\mathit{alias\_graph}$. \\ \hline
%		GetCommonAncestor($\mathit{node_1}$, $\mathit{node_2}$) & Get the 
%		common ancestor of the two nodes $\mathit{node_1}$ and 
%		$\mathit{node_2}$. \\ \hline
%		GetAccessPath($\mathit{node_1}$, $\mathit{node_2}$) & Get the access 
%		path from $\mathit{node_1}$ to $\mathit{node_2}$ \\ \hline
%		\bottomrule[1pt]	
%		\end{tabular}
%	\end{spacing}
%\end{table}

@@

Figure~\ref{fig_pseudocode_field_extract} shows the pseudocode to collect key 
fields that may be protected by a specific lock when are accessed, based on 
lock-set analysis and alias graph. The analysis start analyzing

For each function without a caller function, the analysis

>>>>

Given an alias graph, whether a variable and a lock exist in the same data 
structure can be determined by finding a common ancestor. If they are in the 
same data structure, the lock is likely to protect the variable when it is 
accessed. Figure~\ref{fig_pseudocode_get_access} shows the pseudocode to 
get the field of the accessed variable and the field of the protecting lock in 
the form of access path, if they exist in the same data structure. Given a node 
of an accessed variable and a node of a lock variable, the analysis first gets 
the common ancestor of the two nodes (Line 1). And then, if the common ancestor 
does not exist, the analysis returns a NULL pair (Lines 2-3). Otherwise, the 
analysis gets the access path for the node of the accessed variable and the 
node of the lock variable from the common ancestor (Lines 5-7).

@@

Figure~\ref{fig_pseudocode_field_extract} shows the pseudocode to extract key 
fields that may be protected by a specific lock, based on lock-set analysis and 
the alias graph. The analysis starts from each function without a caller 
function (Lines 16-26) and performs a path-based analysis~\cite{Li:ASPLOS22} 
(Lines 17-25). For each instruction in the code path, the analysis first update 
the alias graph according to the instruction with the four operations (MOVE, 
STORE, LOAD and GEP) (Line 1), and then performs a lock-set analysis (Line 2) 
to get all acquired locks. After updating the alias graph and the lock set, if 
the instruction is a write or a read, the analysis first gets the node of the 
operand with the new alias graph (Lines 4-5). And then, for each node of 
lock variable in the lock set, the analysis uses the alias graph to extract the 
protected data structure field (Lines 6-12). If the field is not NULL, it is 
inserted into the set of key fields (Lines 9-11). For inter-procedural 
analysis, the analysis copies function definition into its call sites and this 
operation is not shown in Figure~\ref{fig_pseudocode_field_extract} for 
convenience.

\PP{S2: Field-access collection.} With the key fields extracted in Stage 1, the 
analysis can only collects accesses to the variables that is protected in some 
cases, because if a variable is never protected by any lock when is accessed, 
it is less likely to be shared by different threads, and thus can not introduce 
no concurrency issue.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=1\linewidth]{figures/fig_pseudocode_access_collect.pdf}
	\figcaption{Pseudocodes of key-field extraction.}
	\label{fig_pseudocode_access_collect}
\end{figure}

Figure~\ref{fig_pseudocode_access_collect} shows the pseudocode to collect all 
accesses to key fields. Similarly to the key-field extraction, this stage also 
performs a path-based analysis. For each instruction in the code path (Lines 
39-45), the analysis first updates the alias graph and the lock set according 
to the handled instruction (Lines 1-2). And then, the access type (either a 
write or a read) is set according to the instruction (Lines 5-8). However, if 
the instruction is not an access operation, the function returns with a NULL 
value for the field access (Line 10). Otherwise, the analysis first gets the 
node of the operand with the new alias graph (Lines 12-13), and then for each 
node of the lock variable in the lock set, the analysis uses the alias graph 
toextract the fields that the accessed variable and the lock variable stored in 
(Lines 16-17). If the fields are found, the field access is returned with the 
found fields (Lines 18-22). If the fields are not found, the accessed variable  
does not protected by any lock. In this case, if the accessed variable is 
stored in a key field, the field access is returned with a NULL lock variable 
(Lines 25-30).

\PP{S3: Locking-rule deduction.} After collecting all accesses to key fields, 
the analysis deduces locking rules based on statistics. However, on the one 
hand, distinguishing different accesses to the same data structure field by 
different program sites is not fine-grained enough, because the access to a 
variable and the acquirement to a specific lock are often packaged in a special 
function, and all accesses under different calling context are regarded as the 
same access. On the other hand, distinguishing different accesses by different 
code paths can also suffer from inaccuracy when a function contains many branch 
statement and cause numerous code paths for the same variable access. Based on 
these consideration, the analysis distinguishes different accesses to the same 
data structure field by different calling contexts. Specifically, given a key 
field {\em f} and a lock {\em l}, the analysis first finds all access to {\em 
f} with lock {\em l} from the collected field accesses in Stage 2, and gets the 
number {\em locked\_access\_num} by counting different calling contexts 
extracted from the code paths of these field accesses. Then, the analysis finds 
all access to {\em f} (no matter which lock is accessed), and gets the number 
{\em all\_access\_num} in the same way as {\em locked\_access\_num}. If the 
ratio {\em locked\_access\_num}/{\em all\_access\_num} is larger than a given 
threshold, and these is at least one write access to {\em f}, the data 
structure field {\em f} is inferred to be protected by the lock {\em l}.
